### 2017-01-10 Simulation_Vignette.Rmd
Started RMarkdown vignette

---
title: "MSAD_PCA_Walkthrough"
author: "Gabriel Odom"
date: "January 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multi-State Adaptive-Dynamic PCA Simulation

We create this package, `mvMonitor`, from the foundation laid by Kazor et al (2016). The `mvMonitor` package is designed to make simulation of multi-state multivariate process monitoring statistics easy and straightforward, as well as streamlining the online process monitoring component.

### The Interval Vector
We begin by simulating an autocorrelated interval-step vector $t$, where the autocorrelated errors of $t$ have initial value $\epsilon_1 \sim \mathcal{U}_{[a,b]}$, where $a = 0.01$ and $b = 2$. This $t$ vector will be sinusoidal with period $\omega = 7 * 24 * 60$ (signifying a weekly period in  minute-level observations) and autocorrelation component $\phi = 0.75$.
```{r t_vec_initialized}
omega <- 60 * 24 * 7
phi <- 0.75
a <- 0.01; b <- 2
mean_t_err <- 0.5 * (a + b) # mean of Uniform(0.01,2)
var_t_err <- (b - a) / 12 # variance of Uniform(0.01,2)

# Create an autocorrelated vector of errors for the random variable t
t_err <- vector(length = omega)
# Initialise the errors so that the overall mean and variance match the sim of
# Kazor et al.
t_err[1] <- rnorm(n = 1,
                  mean = mean_t_err * (1 - phi),
                  sd = sqrt(var_t_err * (1 - phi ^ 2)))
for(s in 2:omega){
  t_err[s] <- phi * t_err[(s - 1)] +
    (1 - phi) * rnorm(n = 1,
                      mean = mean_t_err * (1 - phi),
                      sd = sqrt(var_t_err * (1 - phi ^ 2)))
}

# Now for the vector itself
t_star <- vector(length = omega)
t_star[1] <- -cos(2 * pi / omega) + t_err[1]
for(s in 2:omega){
  t_star[s] <- -cos(2 * pi * s / omega) +
    phi * t_err[(s - 1)] + (1 - phi) * t_err[s]
}
# Now we scale the ts to match the ts from the Unif(0.01,2)
t_star_adj <- ((b - a) * (t_star - min(t_star))) /
  (max(t_star) - min(t_star)) + a
```
Now that we have created the underlying autocorrelated and non-stationary process vector $t$, we perform some simple confirmatory graphing techniques.

```{r t_vec_graphs, echo = FALSE}
plot(t_star_adj,
     type = "l",
     main = "Plot of t Scaled to [0.1,2]",
     xlab = "t")
pacf(t_star,
     main = "Partial ACF for the Series t")
```
```{r t_vec_summary}
mean(t_star); var(t_star)
cor(t_star_adj[2:omega], t_star_adj[1:(omega - 1)])
lmtest::dwtest(t_star_adj[2:omega] ~ t_star_adj[1:(omega - 1)])
```

###  Features and State Features
We will simulate three features, with each feaure operating under $k$ different states. We will use $<x(t), y(t), z(t)>$ instead of the $<x_1(t), x_2(t), x_3(t)>$ notation used in Kazor et al. This will assist us when we introduce feature states. Our state notation will be $<x_k(t), y_k(t), z_k(t)>$ for State $k$.

We create the three features under State 1 (normal operating conditions, or \emph{NOC}) as three functions of $t$:
\begin{align}
\textbf{x}(\textbf{t}) &\equiv \textbf{t} + \boldsymbol\varepsilon_1, \\
\textbf{y}(\textbf{t}) &\equiv \textbf{t} ^ 2 - 3 * \textbf{t} + \boldsymbol\varepsilon_2, \\
\textbf{z}(\textbf{t}) &\equiv -\textbf{t} ^ 3 + 3 * \textbf{t} ^ 2 + \boldsymbol\varepsilon_3,
\end{align}
where $\varepsilon_i \sim N(0, 0.01)$.

```{r state_1_init}
library(scatterplot3d)
library(magrittr)
library(dplyr)

normal_df <- data.frame(t = t_star_adj,
                        err1 = rnorm(n = omega, sd = sqrt(0.01)),
                        err2 = rnorm(n = omega, sd = sqrt(0.01)),
                        err3 = rnorm(n = omega, sd = sqrt(0.01)))

###  State 1  ###
normal_df %<>% mutate(x = t + err1,
                     y = t ^ 2 - 3 * t + err2,
                     z = -t ^ 3 + 3 * t ^ 2 + err3)

# plot(normal_df$t)
# plot(normal_df$x)
# plot(normal_df$y)
# plot(normal_df$z)
# scatterplot3d(x = normal_df$x,
#               y = normal_df$y,
#               z = normal_df$x)
# # It checks out

orig_state_mat <- normal_df %>% select(x,y,z) %>% as.matrix
```

Now we can create data for our other two states. These states will be scaled rotations of the current $<x,y,z>$ set. Thus, we need the `rot_and_scale3d()` function from the `Rotation_Playground.R` file.

```{r state_2_init}
source("inst/Rotation_Playground.R")

###  State 2  ###
state2_angles <- list(yaw = 0,
                      pitch = 90,
                      roll = 30)
state2_scales <- c(1,0.5,2)
state2_mat <- orig_state_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                               scale_factors = state2_scales)
scatterplot3d(state2_mat)

###  State 3  ###
state3_angles <- list(yaw = 90,
                      pitch = 0,
                      roll = -30)
state3_scales <- c(0.25,0.1,0.75)
state3_mat <- orig_state_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                               scale_factors = state3_scales)
scatterplot3d(state3_mat)

###  Combination and Cleaning  ###
# Now concatenate these vectors
normal_df$xState2 <- state2_mat[,1]
normal_df$yState2 <- state2_mat[,2]
normal_df$zState2 <- state2_mat[,3]
normal_df$xState3 <- state3_mat[,1]
normal_df$yState3 <- state3_mat[,2]
normal_df$zState3 <- state3_mat[,3]

# Add a State label column
normal_df %<>%
  mutate(state = rep(rep(c(1, 2, 3), each = 60), (24 / 3) * 7))

# Add a date-time column
normal_df %<>%
  mutate(dateTime = seq.POSIXt(from = as.POSIXct("2016-11-27 00:00:00 CST"),
                               by = "min", length.out = 10080))

###  Hourly Switching Process  ###
state1_df <- normal_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
state2_df <- normal_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
state3_df <- normal_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
normal_switch_df <- bind_rows(state1_df, state2_df, state3_df) %>%
  arrange(dateTime)

normal_switch_df %>%
  filter(state == 3) %>%
  select(x, y, z) %>%
  scatterplot3d(xlim = c(-2, 3),
                ylim = c(-3, 1),
                zlim = c(-1, 5))

normal_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d(xlim = c(-2, 3),
                ylim = c(-3, 1),
                zlim = c(-1, 5))

######  Fault Introduction  ###################################################
faultStart <- 8500

###  Fault 1A  ###
# Shift each <x, y, z> by 2
shift <- 2
fault1A_df <- normal_df %>%
  select(dateTime, state, t, x, y, z) %>%
  mutate(x = x + shift,
         y = y + shift,
         z = z + shift)
fault1A_mat <- fault1A_df %>% select(x,y,z) %>% as.matrix

# State 2
fault1A_S2_mat <- fault1A_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault1A_S2_mat)

# State 3
fault1A_S3_mat <- fault1A_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault1A_S3_mat)

# Now concatenate these vectors
fault1A_df$xState2 <- fault1A_S2_mat[,1]
fault1A_df$yState2 <- fault1A_S2_mat[,2]
fault1A_df$zState2 <- fault1A_S2_mat[,3]
fault1A_df$xState3 <- fault1A_S3_mat[,1]
fault1A_df$yState3 <- fault1A_S3_mat[,2]
fault1A_df$zState3 <- fault1A_S3_mat[,3]

# Hourly switching process
fault1A_state1_df <- fault1A_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault1A_state2_df <- fault1A_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault1A_state3_df <- fault1A_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault1A_switch_df <- bind_rows(fault1A_state1_df,
                               fault1A_state2_df,
                               fault1A_state3_df) %>%
  arrange(dateTime)

fault1A_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

###  Fault 1B  ###
# Shift each x by 2
fault1B_df <- normal_df %>%
  select(dateTime, state, t, x, y, z) %>%
  mutate(x = x + shift)
fault1B_mat <- fault1B_df %>% select(x,y,z) %>% as.matrix

# State 2
fault1B_S2_mat <- fault1B_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault1B_S2_mat)

# State 3
fault1B_S3_mat <- fault1B_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault1B_S3_mat)

# Now concatenate these vectors
fault1B_df$xState2 <- fault1B_S2_mat[,1]
fault1B_df$yState2 <- fault1B_S2_mat[,2]
fault1B_df$zState2 <- fault1B_S2_mat[,3]
fault1B_df$xState3 <- fault1B_S3_mat[,1]
fault1B_df$yState3 <- fault1B_S3_mat[,2]
fault1B_df$zState3 <- fault1B_S3_mat[,3]

# Hourly switching process
fault1B_state1_df <- fault1B_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault1B_state2_df <- fault1B_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault1B_state3_df <- fault1B_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault1B_switch_df <- bind_rows(fault1B_state1_df,
                               fault1B_state2_df,
                               fault1B_state3_df) %>%
  arrange(dateTime)

fault1B_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

###  Fault 2A  ###
# Drift each <x, y, z> by (step - faultStart) / 10 ^ 3
drift_vec <- (1:omega - faultStart) / 10 ^ 3
drift_vec[drift_vec < 0] <- 0
fault2A_df <- normal_df %>%
  select(dateTime, state, t, x, y, z) %>%
  mutate(x = x + drift_vec,
         y = y + drift_vec,
         z = z + drift_vec)
fault2A_mat <- fault2A_df %>% select(x,y,z) %>% as.matrix

# State 2
fault2A_S2_mat <- fault2A_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault2A_S2_mat)

# State 3
fault2A_S3_mat <- fault2A_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault2A_S3_mat)

# Now concatenate these vectors
fault2A_df$xState2 <- fault2A_S2_mat[,1]
fault2A_df$yState2 <- fault2A_S2_mat[,2]
fault2A_df$zState2 <- fault2A_S2_mat[,3]
fault2A_df$xState3 <- fault2A_S3_mat[,1]
fault2A_df$yState3 <- fault2A_S3_mat[,2]
fault2A_df$zState3 <- fault2A_S3_mat[,3]

# Hourly switching process
fault2A_state1_df <- fault2A_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault2A_state2_df <- fault2A_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault2A_state3_df <- fault2A_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault2A_switch_df <- bind_rows(fault2A_state1_df,
                               fault2A_state2_df,
                               fault2A_state3_df) %>%
  arrange(dateTime)

fault2A_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

###  Fault 2B  ###
# Drift each <y, z> by (step - faultStart) / 10 ^ 3
fault2B_df <- normal_df %>%
  select(dateTime, state, t, x, y, z) %>%
  mutate(y = y + drift_vec,
         z = z + drift_vec)
fault2B_mat <- fault2B_df %>% select(x,y,z) %>% as.matrix

# State 2
fault2B_S2_mat <- fault2B_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault2B_S2_mat)

# State 3
fault2B_S3_mat <- fault2B_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault2B_S3_mat)

# Now concatenate these vectors
fault2B_df$xState2 <- fault2B_S2_mat[,1]
fault2B_df$yState2 <- fault2B_S2_mat[,2]
fault2B_df$zState2 <- fault2B_S2_mat[,3]
fault2B_df$xState3 <- fault2B_S3_mat[,1]
fault2B_df$yState3 <- fault2B_S3_mat[,2]
fault2B_df$zState3 <- fault2B_S3_mat[,3]

# Hourly switching process
fault2B_state1_df <- fault2B_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault2B_state2_df <- fault2B_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault2B_state3_df <- fault2B_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault2B_switch_df <- bind_rows(fault2B_state1_df,
                               fault2B_state2_df,
                               fault2B_state3_df) %>%
  arrange(dateTime)

fault2B_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

###  Fault 3A  ###
# Amplify t for each <x, y, z> by 3 * (omega - step) * t / (2 * omega)
amplify_vec <- 2 * (1:omega - faultStart) / (omega - faultStart) + 1
amplify_vec[amplify_vec < 1] <- 1
fault3A_df <- normal_df %>%
  select(dateTime, state, t, err1, err2, err3) %>%
  mutate(t = amplify_vec * t) %>%
  mutate(x = t + err1,
         y = t ^ 2 - 3 * t + err2,
         z = -t ^ 3 + 3 * t ^ 2 + err3)
fault3A_mat <- fault3A_df %>% select(x,y,z) %>% as.matrix

# State 2
fault3A_S2_mat <- fault3A_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault3A_S2_mat)

# State 3
fault3A_S3_mat <- fault3A_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault3A_S3_mat)

# Now concatenate these vectors
fault3A_df$xState2 <- fault3A_S2_mat[,1]
fault3A_df$yState2 <- fault3A_S2_mat[,2]
fault3A_df$zState2 <- fault3A_S2_mat[,3]
fault3A_df$xState3 <- fault3A_S3_mat[,1]
fault3A_df$yState3 <- fault3A_S3_mat[,2]
fault3A_df$zState3 <- fault3A_S3_mat[,3]

# Hourly switching process
fault3A_state1_df <- fault3A_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault3A_state2_df <- fault3A_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault3A_state3_df <- fault3A_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault3A_switch_df <- bind_rows(fault3A_state1_df,
                               fault3A_state2_df,
                               fault3A_state3_df) %>%
  arrange(dateTime)

fault3A_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

###  Fault 3B  ###
# Dampen t for z by log|t|
t_log <- normal_df$t
t_log[faultStart:omega] <- log(t_log[faultStart:omega])
fault3B_df <- normal_df %>%
  select(dateTime, state, t, err1, err2, err3) %>%
  mutate(t_damp = t_log) %>%
  mutate(x = t + err1,
         y = t ^ 2 - 3 * t + err2,
         z = -t_damp ^ 3 + 3 * t_damp ^ 2 + err3)
fault3B_mat <- fault3B_df %>% select(x,y,z) %>% as.matrix

# State 2
fault3B_S2_mat <- fault3B_mat %*% rot_and_scale3d(rot_angles = state2_angles,
                                                  scale_factors = state2_scales)
scatterplot3d(fault3B_S2_mat)

# State 3
fault3B_S3_mat <- fault3B_mat %*% rot_and_scale3d(rot_angles = state3_angles,
                                                  scale_factors = state3_scales)
scatterplot3d(fault3B_S3_mat)

# Now concatenate these vectors
fault3B_df$xState2 <- fault3B_S2_mat[,1]
fault3B_df$yState2 <- fault3B_S2_mat[,2]
fault3B_df$zState2 <- fault3B_S2_mat[,3]
fault3B_df$xState3 <- fault3B_S3_mat[,1]
fault3B_df$yState3 <- fault3B_S3_mat[,2]
fault3B_df$zState3 <- fault3B_S3_mat[,3]

# Hourly switching process
fault3B_state1_df <- fault3B_df %>%
  filter(state == 1) %>%
  select(dateTime, state, x, y, z)
fault3B_state2_df <- fault3B_df %>%
  filter(state == 2) %>%
  select(dateTime, state, x = xState2, y = yState2, z = zState2)
fault3B_state3_df <- fault3B_df %>%
  filter(state == 3) %>%
  select(dateTime, state, x = xState3, y = yState3, z = zState3)
fault3B_switch_df <- bind_rows(fault3B_state1_df,
                               fault3B_state2_df,
                               fault3B_state3_df) %>%
  arrange(dateTime)

fault3B_switch_df %>%
  select(x, y, z) %>%
  scatterplot3d()

######  The Seven Data Frames  ################################################
# The normal_df data frame is without fault. We will remove rows faultStart to
# omega, and replace these rows with the same rows from each each of the fault
# data frames.
# Turn all the data frames into xts objects for Kazor's function to work.
# NOTE - this follows a depreciated version of zoo, so we need the dateTime
# column as the rownames of the data frame, not as an index.
# EDITED NOTE - B. Barnard and I rebuilt Kazor's code as a functional package.
# The functions in these packages use xts matrices as inputs.
library(xts)
normal_switch_xts <- xts(normal_switch_df[,-1],
                         order.by = normal_switch_df[,1])

# Fault 1A
normal_w_fault1A_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                              fault1A_switch_df[faultStart:omega,])
fault1A_xts <- xts(normal_w_fault1A_df[,-1], order.by = normal_switch_df[,1])

# Fault 1B
normal_w_fault1B_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                                 fault1B_switch_df[faultStart:omega,])
fault1B_xts <- xts(normal_w_fault1B_df[,-1], order.by = normal_switch_df[,1])

# Fault 2A
normal_w_fault2A_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                                 fault2A_switch_df[faultStart:omega,])
fault2A_xts <- xts(normal_w_fault2A_df[,-1], order.by = normal_switch_df[,1])

# Fault 2B
normal_w_fault2B_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                                 fault2B_switch_df[faultStart:omega,])
fault2B_xts <- xts(normal_w_fault2B_df[,-1], order.by = normal_switch_df[,1])

# Fault 3A
normal_w_fault3A_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                                 fault3A_switch_df[faultStart:omega,])
fault3A_xts <- xts(normal_w_fault3A_df[,-1], order.by = normal_switch_df[,1])

# Fault 3B
normal_w_fault3B_df <- bind_rows(normal_switch_df[1:(faultStart - 1),],
                                 fault3B_switch_df[faultStart:omega,])
fault3B_xts <- xts(normal_w_fault3B_df[,-1], order.by = normal_switch_df[,1])

faults_ls <- list(normal = normal_switch_xts,
                  fault1A = fault1A_xts,
                  fault1B = fault1B_xts,
                  fault2A = fault2A_xts,
                  fault2B = fault2B_xts,
                  fault3A = fault3A_xts,
                  fault3B = fault3B_xts)

# Double check that faults are visible and according to plan:
# faults_ls$fault1A %>% select(x) %>% plot.ts()
# faults_ls$fault1A %>% select(y) %>% plot.ts()
# faults_ls$fault1A %>% select(z) %>% plot.ts()
# faults_ls$fault1B %>% select(x) %>% plot.ts()
# faults_ls$fault1B %>% select(y) %>% plot.ts()
# faults_ls$fault1B %>% select(z) %>% plot.ts()
# faults_ls$fault2A %>% select(x) %>% plot.ts()
# faults_ls$fault2A %>% select(y) %>% plot.ts()
# faults_ls$fault2A %>% select(z) %>% plot.ts()
# faults_ls$fault2B %>% select(x) %>% plot.ts()
# faults_ls$fault2B %>% select(y) %>% plot.ts()
# faults_ls$fault2B %>% select(z) %>% plot.ts()
# faults_ls$fault3A %>% select(x) %>% plot.ts()
# faults_ls$fault3A %>% select(y) %>% plot.ts()
# faults_ls$fault3A %>% select(z) %>% plot.ts()
# faults_ls$fault3B %>% select(x) %>% plot.ts()
# faults_ls$fault3B %>% select(y) %>% plot.ts()
# faults_ls$fault3B %>% select(z) %>% plot.ts()
faults_ls$fault1A$x %>% plot.xts()
faults_ls$fault1A$y %>% plot.xts()
faults_ls$fault1A$z %>% plot.xts()
faults_ls$fault1B$x %>% plot.xts()
faults_ls$fault1B$y %>% plot.xts()
faults_ls$fault1B$z %>% plot.xts()
faults_ls$fault2A$x %>% plot.xts()
faults_ls$fault2A$y %>% plot.xts()
faults_ls$fault2A$z %>% plot.xts()
faults_ls$fault2B$x %>% plot.xts()
faults_ls$fault2B$y %>% plot.xts()
faults_ls$fault2B$z %>% plot.xts()
faults_ls$fault3A$x %>% plot.xts()
faults_ls$fault3A$y %>% plot.xts()
faults_ls$fault3A$z %>% plot.xts()
faults_ls$fault3B$x %>% plot.xts()
faults_ls$fault3B$y %>% plot.xts()
faults_ls$fault3B$z %>% plot.xts()

######  MSAD-PCA Implementation  ##############################################
# Sys.time()
# results_ls <- mspMonitor(data = faults_ls$normal[,2:4],
#                          labelVector = faults_ls$normal[,1],
#                          trainObs = 2880,
#                          updateFreq = 1440)
# Sys.time() # 14 seconds for 720 train, 360 update; 5 seconds for 1440, 720;
# # 1 second for 2880 train (two days), and 1440 update (one day).
# 
# falseAlarmRate_SPE <- sum(results_ls$Alarms[,1] == "Alarm") /
#   nrow(results_ls$Alarms)
# falseAlarmRate_T2 <- sum(results_ls$Alarms[,2] == "Alarm") /
#   nrow(results_ls$Alarms)
# 
# faultTime <- zoo::index(faults_ls$normal[faultStart,])
# # Why don't we have 10080 - 2880 rows in the Alarm matrix??
# str(results_ls$Alarms)

```